### 2022-08-11 draft.tex
Deploying to gh-pages from @ FrancisBanville/ms_maxent_networks@b53464e1b97a8e6f7721c7453fb60da2b54765ad 🚀

The degree distribution \(p(k)\) represents the probability that a
species has \(k\) interactions in a food web, with
\(k = k_{in} + k_{out}\). It can thus be directly obtained from the
joint degree distribution:

@@

In fig.~S2, we show that the degree distribution of maximum entropy,
given \(S\) and \(L\), predicts very low probabilities that a species
will be isolated in its food web (i.e.~having \(k = 0\)). As MacDonald,
Banville, and Poisot (2020) pointed out, the size of food webs should at
least be of \(S-1\) interactions, since a lower number would yield
isolated species, i.e.~species without any predators or preys. Our
results show that, under our purely information-theoretic model, the
probability that a species is isolated is quite high when the total
number of interactions is below this threshold. The expected proportion
of isolated species rapidly declines by orders of magnitude with
increasing numbers of species and interactions.

The degree distribution could also have been obtained directly using the
principle of maximum entropy, as discussed in Williams (2011). This
gives the following distribution:

@@

This can be solved numerically using the constraint of the average
degree \(\langle k \rangle = \frac{2L}{S}\) of a species. Note that the
mean degree is twice the value of the linkage density, because every
link must be counted twice when we add in and out-degrees together.

@@

The numerical solution is identical to the one we obtained using the
joint degree distribution as an intermediate. Ecologists wanting to
model a system without considering isolated species could simply change
the lower limit of \(k\) to \(1\) and solve the resulting equation
numerically.

In this section, we showed how important measures of food-web structure,
namely the degree distribution and the joint degree distribution, could
be derived with the principle of maximum entropy using minimal knowledge
on a biological community. This type of models, although useful to make
least-biased predictions on many network properties, can be hard to
apply for other measures. Indeed, there are dozens of measures of
network structure (Delmas et al. 2019) and many are not calculated with
mathematical equations, but with algorithms. Moreover, the applicability
of this method to empirical systems is limited by the state variables we
can actually measure and use. In the next section, we propose a more
flexible method to predict many measures of network structure
simultaneously, i.e.~by finding heuristically the configuration of the
network having maximum entropy given partial knowledge of its emerging
structure.

>>>>

The joint degree distribution derived using MaxEnt
(eq.~\ref{eq:lagrange3_jdd}) can be used to obtain the degree
distribution of maximum entropy. Indeed, the degree distribution
\(p(k)\) represents the probability that a species has \(k\)
interactions in its food web, with \(k = k_{in} + k_{out}\). It can thus
be obtained from the joint degree distribution as follows:

@@

The degree distribution could have also been obtained directly using the
principle of maximum entropy, as discussed in Williams (2011). This
gives the following distribution:

@@

This can be solved numerically using the constraint of the average
degree \(\langle k \rangle = \frac{2L}{S}\) of a species, yielding an
identical solution to the one obtained using the joint degree
distribution as an intermediate. Note that the mean degree is twice the
value of the linkage density, because every link must be counted twice
when we add in and out-degrees together.

@@

One aspect of the degree distribution that informs us of its ecological
realism is the number of isolated species it predicts. As MacDonald,
Banville, and Poisot (2020) pointed out, the size of food webs should at
least be of \(S-1\) interactions, since a lower number would yield
isolated species, i.e.~species without any predators or preys. Because
non-basal species must eat to survive, isolated species could indicate
that other species are missing or they could simply be removed from the
food web. In fig.~S2, we show that the degree distribution of maximum
entropy, given \(S\) and \(L\), gives very low probabilities that a
species will be isolated in its food web (i.e.~having \(k = 0\)) above
the \(S-1\) threshold. However, under our purely information-theoretic
model, the probability that a species is isolated is quite high when the
total number of interactions is below \(S-1\). Moreover, the expected
proportion of isolated species rapidly declines by orders of magnitude
with increasing numbers of species and interactions (fig.~S2). This
supports the ecological realism of the degree distribution of maximum
entropy derived above. Nevertheless, ecologists wanting to model a
system without allowing isolated species could simply change the lower
limit of \(k\) to \(1\) in eq.~\ref{eq:lagrange2_dd} and solve the
resulting equation numerically.

Using the analytical approach, we showed how important measures of
food-web structure, namely the degree distribution and the joint degree
distribution, could be derived with the principle of maximum entropy
given minimal knowledge about a biological community. This type of
models, although useful to make least-biased predictions on many network
properties, can be hard to apply for other measures. Indeed, there are
dozens of measures of network structure (Delmas et al. 2019) and many
are not calculated with mathematical equations, but with algorithms.
Moreover, the applicability of this method to empirical systems is
limited by the state variables we can actually measure and use. In the
next section, we propose a more flexible method to predict many measures
of network structure simultaneously, i.e.~by finding heuristically the
network configuration having maximum entropy given partial knowledge of
its emerging structure.
